\LARGE \bf Cross Validation of Emotional Features with a non-Anthropomorphic Platform

### 2017-01-09 RO-MAN-2017-1/Parts/Antecedents.tex
First version of RO-Man article base on the last case study. this article has been improved from the one submitted to HRI, but it is still necessary to review the general idea and how to make everything more clear

An experiment was done to assess precise values for linear velocity, angular velocity, oscillation angle, direction and orientation that could be used to express happiness, angry, fear and sadness, which correspond to four basic emotions suggested by Ekman~\cite{Ekman2001}. These features were selected after study emotion projection in robotics, humans (i.e. previous work and theatre) and previous case studies. The robotic platform used in the experiment is holonomic, which are characterized by the possibility to move in any direction without necessity to have a specific orientation, i.e., they are free to move taking any desired orientation. The experiment was performed at the university campus during the months of June and July of 2015.  A total of 49 volunteers were involved: 12 female and 37 male. The average age of the participants was 25.28 with standard deviation of 2.8, with a minimum age of 20 and maximum of 32.

A Google form was used to collect participants' answers. The form included the four emotions plus two mental states (i.e. excitement and tenderness) and option ''other''. These mental state were added to see if people could confuse the desired emotion with them. Each participant was exposed to twenty over one hundred and ninety five possible treatments. The whole process, including a brief explanation and assessment lasted  from 10 to 15 minutes. This was decided because each subject was a volunteer and would not perceive any monetary remuneration, so the time dedicated to the experiment had to be kept limited. The twenty treatments were selected picking a number without replacement from 1 to 195. For each presented treatment, participants were asked to give an intensity perceived for each option in the form.

After collecting all the data, it was created a table for each treatment. Each table contain the following information mean, standard deviation, and median. ANOVA test was not possible to be used over the data because the assumption of normality is not achieved in the collected data. This was check using the Shapiro-Wilk Test.
Additional to these table, a contingency table for each emotion was generated in each treatment as it is depicted in  Table~\ref{table:table_contingency},  where the intensity for the other emotions is calculated as the mean of them, including the option of ''other''. For all tables, including the contingency, were calculated the Krippendorff's alpha agreement~\cite{Krippendorff2007} ($\alpha$), which is a reliability coefficient to measure the agreement among different participants. Unlike other coefficients (Kappa), $\alpha$ is a generalization of several known reliability indices, and it applies to:

@@

This calculation was done using the R package \textit{irr}. To improve the table interpretation, it was decided to just record the emotions' alpha values that had a mean greater than zero. Therefore, tables with a top ten ranking have been set up. The raking considered: (i) the mean of the respective emotions, (ii) the alpha agreement for the respective emotion, and (iii) the alpha agreement for the treatment. The decision to give more importance to emotion's alpha rather than intensity average was taken basing on the consideration that most participants agreed on their observation. From the results was possible to notice:

>>>>

An experiment was done to assess precise values for linear velocity, angular velocity, oscillation angle, direction and orientation that could be used to express happiness, angry, fear and sadness, which correspond to four of six basic emotions suggested by Ekman~\cite{Ekman2001}. These features were selected after study emotion projection in robotics, humans (i.e. previous work and theatre) and previous case studies. The robotic platform used in the experiment is holonomic, which are characterized by the possibility to move in any direction without necessity to have a specific orientation, i.e., they are free to move taking any desired orientation. The experiment was performed at the university campus during the months of June and July of 2015.  A total of 49 volunteers were involved: 12 female and 37 male. The average age of the participants was 25.28 with standard deviation of 2.8, with a minimum age of 20 and maximum of 32.

A Google form was used to collect participants' answers. The form included four emotions plus two mental states (i.e. excitement and tenderness) and option ''other''. These mental states were added to see if people could confuse the desired emotion with them. Each participant was exposed to twenty over one hundred and ninety five possible treatments. The whole process, including a brief explanation and assessment lasted  from 10 to 15 minutes. This was decided because each subject was a volunteer and would not perceive any monetary remuneration, so the time dedicated to the experiment had to be kept limited. The twenty treatments were selected picking a number without replacement from 1 to 195. For each presented treatment, participants were asked to give an intensity perceived for each option in the form.

After collecting all the data, it was created a table for each treatment. Each table contained the following information mean, standard deviation, and median. ANOVA test was not possible to be used over the data because the assumption of normality is not achieved in the collected data. This was check using the Shapiro-Wilk Test.
Additional to these table, a contingency table for each emotion was generated in each treatment as it is depicted in  Table~\ref{table:table_contingency},  where the intensity for the other emotions is calculated as the mean of them, including the option of ''other''. For all tables, including the contingency, were calculated the Krippendorff's alpha agreement~\cite{Krippendorff2007} ($\alpha$), which is a reliability coefficient to measure the agreement among different participants. Unlike other coefficients (Kappa), $\alpha$ is a generalization of several known reliability indices, and it applies to:

@@

This calculation was done using the R package \textit{irr}. To improve table interpretation, it was decided to just record the emotions' alpha values that had a mean greater than zero. Therefore, tables with a top ten ranking have been set up. The raking considered: (i) the mean of the respective emotions, (ii) the alpha agreement for the respective emotion, and (iii) the alpha agreement for the treatment. The decision to give more importance to emotion's alpha rather than intensity average was taken basing on the consideration that most participants agreed on their observation. From the results was possible to notice:

### 2017-01-09 RO-MAN-2017-1/Parts/CaseStudy.tex
First version of RO-Man article base on the last case study. this article has been improved from the one submitted to HRI, but it is still necessary to review the general idea and how to make everything more clear

The case study was done at Researcher's Night 2015 with two main objectives (i) cross-validate the findings obtained from the experiment, and (ii) use the Emotional Enrichment System to verify whether the participants would prefer scenes when the robot expresses emotions or rather moves without any emotion expression. The Emotional Enrichment System was used in both cases. To reduce scene variability was used two web-cams and eight Alvar tags to informed a Kalman filter to improve robots localization in the stage. The detection of the AR tags was done through the use of the ROS package ar\_track\_alvar~\cite{artag2015}. The distribution of the web-cams and the tags are depicted in Figure~\ref{fig:setup_fourth}.

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{./Images/FourthCase.png} 
	\caption{Environment setup for the case study.}
	\label{fig:setup_fourth}
\end{figure}

@@

The parameters selected for each of four emotions (\textit{Anger}, \textit{Happiness}, \textit{Sadness} and \textit{Fear}) are shown in Table~\ref{table:selected_fourth}. The main considerations to select the two implementations for each emotion were: (i) the linear velocity should be greater than $0$. In other words the robot should show some linear displacement. And (ii) it should be in the top 10 list of the emotion obtained in the experiment.

\begin{table}[h]
\centering
\small
\caption{Parameters' values selected from the experiment.}
		\label{table:selected_fourth}
		\begin{tabular}{|c|p{0.9 cm}|p{0.9 cm}|p{0.9 cm}|p{1.05 cm}|p{0.9 cm}|}
			\hline
%\rotatebox{90}{\textbf{Emotion } }&
%\rotatebox{90}{\textbf{Direction  ($rad$)}}&
%\rotatebox{90}{\textbf{Orientation ($rad$)} }&
%\rotatebox{90}{\textbf{Linear Velocity ($mm/s$) }}&
%\rotatebox{90}{\textbf{Angular Velocity ($rad/s$) }}&
%\rotatebox{90}{\textbf{Angle ($rad$)}}\\	
\textbf{Emotion}&\textbf{Direc-tion  ($rad$)} & \textbf{Orien-tation ($rad$)} & \textbf{Linear Velocity ($mm/s$) } & \textbf{Angular Velocity ($rad/s$) } & \textbf{Angle ($rad$)} \\
			\hline
			Happiness 1&$0$&$0$&$500$&$3$&$0.349$\\
			\hline
			Happiness 2&$0$&$0$&$900$&$3$&$0.174$\\
			\hline
			Anger 1&$\pi$&$0$&$500$&$3$&$0.087$\\
			\hline
			Anger 2&$0$&$0$&$900$&$1$&$0.087$\\
			\hline
			Fear 1&$\pi$&$\pi$&$900$&$2$&$0.174$\\
			\hline
			Fear 2&$\pi$&$\pi$&$500$&$2$&$0.087$\\
			\hline
			Sadness 1&$\pi$&$0$&$200$&$1$&$0.349$\\
			\hline
			Sadness 1&$0$&$\pi$&$200$&$1$&$0.349$\\
			\hline
			\end{tabular}
\end{table}

@@

Stage discretization was used to give zones of movements instead of absolute positions. This idea was brought from human theatrical actors, who prepared their movements based on zones on the stage~\cite{wilson2009theatre}. This allows them to adapt their position based on other actors. Therefore, the stage was discretized in 9x9 matrix as is shown in Figure~\ref{fig:stage_division}. Robot's movements are given in terms of the matrix positions to the Emotional Enrichment System. This allows the adaptation to different stage dimensions because robot's final position is calculated by the Emotional Enrichment System during execution, which takes under consideration the stage dimensions given. For instance, during the scene's preparation in the laboratory the stage was 3 meters per 3 meters, but in the final presentation the stage was 2.5 meters per 2.5 meters.

@@

The scene's description is the following: the robot starts in the middle of the stage to move to the upstage right (See~\cite{Musical}), close to the right wing. Then, the robot moves to upstage right center and rotates by $\pi/2$ left (See~\cite{Artopia}). Next the robot moves to the right center to then go to the center. When it arrives there, it turns full back and move backwards to downstage center with a full front orientation. There, it turns full back to move to center. Finally the robot turns to profile right and it does a step back; then it goes to the upstage center and then upstage right. The sequence of movements programmed to the robot are depicted in Figure~\ref{fig:movement}.
\begin{figure*}
	\centering
	\includegraphics[width=0.95\textwidth]{./Images/fourthCaseSceneD.png} 
	\caption{Sequence of movements done by the robot. The red arrows show the trajectory done by the robot, while the numbers show the order among the movements. a) The first ten movements b) The last five movements }
	\label{fig:movement}
\end{figure*}

The relation between emotion and movement is as follow: movements one, two, three, four and five are expressed without any emotion. Movements six, seven, eight, nine and ten show fear. Movement eleven depicts happiness, and the remaining movements depict sadness. The two scenes are executed by the Emotional Enrichment System using the same ''script'' and the emotion selection is done manually via graphical interface (Figure~\ref{fig:graphical_interface}).

\begin{figure}
	\centering
	\includegraphics[width=0.48\textwidth]{./Images/InterfaceExperiment.png} 
	\caption{Graphical interface used to communicate with the Emotional Enrichment System. a) It is the interface used to send the actions sequence. b) It is the interface used to send a emotion and its intensity.}
	\label{fig:graphical_interface}
\end{figure}

@@

This case study was done during Researchers' Night, 2015. During a period of two days, people were asked to participate to this study. Each subject was exposed to two rounds, in each one the robot was performing a different emotion. And  they were also exposed twice to a small scene, one with emotion and other without emotions. The emotions showed in each trial and the order of the scenes (with or without emotion) were generated randomly beforehand. The total number of volunteers was 256: 128 males, 126 females, and 2 that chose not to specify their gender. The average age was 27.29 years, with standard deviation of 16.58, minimum age was 4 and maximum 76.

>>>>

The case study was done at Researcher's Night 2015 with two main objectives:
\begin{enumerate}
	\item Cross-validate the findings obtained from a previous experiment~\cite{Angel2017-2}. This experiment was done to study the attribution of different linear and angular velocities, oscillation angel, direction and orientation of the platform to \textit{Anger}, \textit{Happiness}, \textit{Sadness} and \textit{Fear}. A total of 196 were designed, but just 20 treatments were presented to each participant. The questionnaire provided included the four emotions and two mental states (i.e. Excitement and tenderness).
	\item Verify if participants would prefer scenes when the robot expresses emotions or rather moves without any emotion expression. 
\end{enumerate}
To reduce scene variability was used two web-cams and eight Alvar tags to informed a Kalman filter to improve robots localization in the stage. AR tags were detected using the ROS package ar\_track\_alvar~\cite{artag2015}. The distribution of the web-cams and the tags are depicted in Figure~\ref{fig:setup_fourth}.

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{./Images/FourthCase.png} 
	\caption{Environment setup for the case study. The crosses represent the starting points.}
	\label{fig:setup_fourth}
\end{figure}

@@

The parameters selected from the experiment to express \textit{Anger}, \textit{Happiness}, \textit{Sadness} and \textit{Fear} are shown in Table~\ref{table:selected_fourth}. As it could be observed, two implementations of each emotions were selected. These implementations were selected considering: (i) the linear velocity should be greater than $0$. In other words the robot should show some linear displacement. And (ii) it should be in the top 10 list of the emotion obtained in the experiment.

\begin{table}
\centering
\small
\caption{Parameters' values selected from the experiment.}
		\label{table:selected_fourth}
		\begin{tabular}{|c|p{0.9 cm}|p{0.9 cm}|p{0.9 cm}|p{1.05 cm}|p{0.9 cm}|}
			\hline
%\rotatebox{90}{\textbf{Emotion } }&
%\rotatebox{90}{\textbf{Direction  ($rad$)}}&
%\rotatebox{90}{\textbf{Orientation ($rad$)} }&
%\rotatebox{90}{\textbf{Linear Velocity ($mm/s$) }}&
%\rotatebox{90}{\textbf{Angular Velocity ($rad/s$) }}&
%\rotatebox{90}{\textbf{Angle ($rad$)}}\\	
\textbf{Emotion}&\textbf{Direc-tion  ($rad$)} & \textbf{Orien-tation ($rad$)} & \textbf{Linear Velocity ($mm/s$) } & \textbf{Angular Velocity ($rad/s$) } & \textbf{Angle ($rad$)} \\
			\hline
			Happiness-1&$0$&$0$&$500$&$3$&$0.349$\\
			\hline
			\co Happiness-2&\co $0$&\co $0$&\co $900$&\co $3$&\co $0.174$\\
			\hline
			Anger-1&$\pi$&$0$&$500$&$3$&$0.087$\\
			\hline
			\co Anger-2&\co $0$&\co $0$&\co $900$&\co $1$&\co $0.087$\\
			\hline
			Fear-1&$\pi$&$\pi$&$900$&$2$&$0.174$\\
			\hline
			\co Fear-2&\co $\pi$&\co $\pi$&\co $500$&\co $2$&\co $0.087$\\
			\hline
			Sadness-1&$\pi$&$0$&$200$&$1$&$0.349$\\
			\hline
			\co Sadness-2&\co $0$&\co $\pi$&\co $200$&\co $1$&\co $0.349$\\
			\hline
			\end{tabular}
\end{table}

@@

Stage discretization was used to give zones of movements instead of absolute positions. This idea was brought from human theatrical actors, who prepared their movements based on zones on the stage~\cite{wilson2009theatre}. This allows them to adapt their position based on other actors and stage dimensions. Therefore, the stage was discretized in 9x9 matrix as is shown in Figure~\ref{fig:stage_division}. Robot's movements are given in terms of the matrix positions to the Emotional Enrichment System. Robot's final position is calculated by the Emotional Enrichment System during execution. For instance, during the scene's preparation stage was 3 meters per 3 meters, but in the final presentation it was 2.5 meters per 2.5 meters.

@@

Scene's description is the following: the robot starts in the middle of the stage to move to the upstage right (See~\cite{Musical}), close to the right wing. Then, the robot moves to upstage right center and rotates to $\pi/2$ left (See~\cite{Artopia}). Next the robot moves to the right center. Then it goes to the center. When it arrives there, it turns full back and move backwards to downstage center with a full front orientation. There, it turns full back to move to center. Finally the robot turns to profile right and it does a step back; then it goes to the upstage center and then upstage right. The sequence of movements programmed to the robot are depicted in Figure~\ref{fig:movement}.
\begin{figure*}
	\centering
	\includegraphics[width=0.95\textwidth]{./Images/fourthCaseSceneD.png} 
	\caption{Sequence of movements provided to the robot. The red arrows show the trajectory, while the numbers show the order among the movements. a) The first ten movements b) The last five movements }
	\label{fig:movement}
\end{figure*}

The relation between emotion and movement is as follow: movements one to five do not express any emotion. Movements six to ten show fear. Movement eleven depicts happiness, and the remaining movements depict sadness. The actions describing this scene are executed by the Emotional Enrichment System and the emotion selection is done manually via a graphical interface.

@@

This case study was done during Researchers' Night, 2015. During a period of two days, people were asked to participate to this study, which was divided in to parts. In the first one, each subject was exposed to two rounds of emotions. The emotion presented and the order were generated randomly before hand. In the second part, they were explained that a small scene was going to be presented twice. Thus, they should to selected the one they like it more. The order of the scenes (e.g. with or without emotion) were generated beforehand. The total number of volunteers was 256: 128 males, 126 females, and 2 that chose not to specify their gender. The average age was 27.29 years, with standard deviation of 16.58, minimum age was 4 and maximum 76.

### 2017-01-09 RO-MAN-2017-1/Parts/Conclusions.tex
First version of RO-Man article base on the last case study. this article has been improved from the one submitted to HRI, but it is still necessary to review the general idea and how to make everything more clear

The case study presented in this paper was done to cross validate the findings in the experiment and verify whether the participants would prefer scenes when the robot expresses emotions or rather moves without any emotion expression. For each one of four emotions (i.e. Anger, Happiness, Sadness and Fear) studied in the experiment were selected a two set of parameters. These parameters were described in files and used as input to the Enrichment Emotional System, which was in charge to combine actions with emotions. The results show that both implementations of happiness were confused with anger and excitement, while one implementation of anger was just confused with excitement. Both implementations of sadness were confused with tenderness and fear. Both implementations of fear had a recognition rate over 50\%. Scene's results show that people prefer scenes with emotional movements and there is not any difference in gender.

>>>>

The case study presented in this paper was done to cross validate the findings in the experiment and verify whether the participants would prefer scenes when the robot expresses emotions or rather moves without any emotion expression. For each one of four emotions (i.e. Anger, Happiness, Sadness and Fear) studied in the experiment were selected a two set of parameters.The results show that both implementations of happiness were confused with anger and excitement, while one implementation of anger was just confused with excitement. Both implementations of sadness were confused with tenderness and fear. Both implementations of fear had a recognition rate over 50\%. Scenes results show that people prefer scenes with emotional movements and there is not any difference in gender.

### 2017-01-09 RO-MAN-2017-1/Parts/Introduction.tex
First version of RO-Man article base on the last case study. this article has been improved from the one submitted to HRI, but it is still necessary to review the general idea and how to make everything more clear

Emotions and mental states are not just expressed through facial expression but also by body through postures and other features~\cite{Gelder2008}. Many psychological studies have been focused on understanding the role of human face in emotion projection~\cite{Ekman2004},~\cite{kleinsmith2012affective} and mental states. This trend has been followed by the robotics community, where anthropomorphic faces (e.g.,~\cite{Arras2012},~\cite{Breazeal2002}) and bodies (e.g.,~\cite{Canamero2010},~\cite{haering2011},~\cite{Destephe2013}) have been used to convey emotions.
However in many situations presence of anthropomorphic elements would be out of place and not justified by the main robot's functionalities. Most of the current and future robotics platforms on the market will not require anthropomorphic faces~\cite{Breazeal2002} or limbs~\cite{Li2011}, for instance in floor cleaning robots, anthropomorphic characteristics could even be detrimental to robot's task accomplishment.
This generates the necessity to study other mechanisms that could help to project emotions, which could give people an idea about the robots' state, and engage the user in long term relations.

It has been noticed that the amount of works studying non-anthropomorphic features in robotics (e.g.,~\cite{Saerbeck2010,Lakatos2014,Sharma2013,Novika2015}) remains still small in comparison to those exploiting anthropomorphic features. Moreover, these works do not give specific range of values for the characteristics used to expressed  emotions implemented. For example Suk and collaborators~\cite{NAM2014} in their study give specific values for acceleration and curvature,  but their connection to specific emotions is not explicited. Rather, they establish a relationship between their values and pleasure/arousal dimensions. This could help to select specific features to convey certain emotions, but it would be even better to know the precise range of values to be assigned to each characteristics to avoid  misinterpretation with wrong emotions. These values could be used in social robotics to show emotions and as consequence increase their acceptance.

This paper is organized as follows. Section~\ref{sec:related_work} introduce some previous works done in conveying emotion. Then the experiment is briefly described in section~\ref{sec:experiment}. Section~\ref{sec:system} presents the platform used in the case study and the Emotional Enrichment System. Finally, sections~\ref{sec:case} and~\ref{sec:results} presented case study's design and results.

>>>>

Emotions and mental states are not just expressed through facial expression but also by body through postures and other features~\cite{Gelder2008}. Many psychological studies have been focused on understanding the role of human face in emotion projection~\cite{Ekman2004},~\cite{kleinsmith2012affective} and mental states. This trend has been followed by the robotics community, where anthropomorphic faces (e.g.,~\cite{Arras2012},~\cite{Breazeal2002}) and bodies (e.g.,~\cite{Canamero2010},
%~\cite{haering2011},
~\cite{Destephe2013}
) have been used to convey emotions.
However in many situations presence of anthropomorphic elements would be out of place and not justified by the main robot's functionalities. Most of the current and future robotics platforms on the market will not require anthropomorphic faces~\cite{Breazeal2002} or limbs~\cite{Li2011}, for instance in floor cleaning robots, anthropomorphic characteristics could even be detrimental to robot's task accomplishment.
This generates the necessity to study other mechanisms that could help to project emotions, which could give people an idea about the robots' state, and engage the user in long term relations.

It has been noticed that the amount of works studying non-anthropomorphic features in robotics (e.g.,~\cite{Saerbeck2010}
%,Lakatos2014,Sharma2013,Novika2015}
) remains still small in comparison to those exploiting anthropomorphic features. Moreover, these works do not give specific range of values for the characteristics used to expressed  emotions implemented. For example Suk and collaborators~\cite{NAM2014} in their study give specific values for acceleration and curvature,  but their connection to specific emotions is not explicited. Rather, they establish a relationship between their values and pleasure/arousal dimensions. This could help to select specific features to convey certain emotions, but it would be even better to know the precise range of values to be assigned to each characteristics to avoid  misinterpretation with wrong emotions. These values could be used in social robotics to show emotions and as consequence increase their acceptance.

This paper is organized as follows. Section~\ref{sec:related_work} introduce some previous works done in conveying emotion. Section~\ref{sec:system} presents the platform used in the case study and the Emotional Enrichment System. Finally, sections~\ref{sec:case} and~\ref{sec:results} presented case study's design and results.

### 2017-01-09 RO-MAN-2017-1/Parts/RelatedWork.tex
First version of RO-Man article base on the last case study. this article has been improved from the one submitted to HRI, but it is still necessary to review the general idea and how to make everything more clear

Although it is not possible to apply a direct mapping from human studies to robots~\cite{Saerbeck2007,Canamero2010}, human studies provide guidelines about possible features and values that could be used to generate emotional motion.

\subsection{Human Studies}

Emotion plays an important role in human-human interaction, and can be expressed through diverse channels such as body gestures and poses, body movements, face expressions. Given the complex structure of human face, where more than 43 muscles act, face configurations are considered as a primary channel to express emotion. As a consequence, many works have focused on facial expression, mainly, but not only, influenced by the work done by Ekman~\cite{Ekman2004}. However, the role that body plays in emotion projection was recognized, too, and started to be studied~\cite{Gelder2008,Wallboot1998}. Nevertheless, the amount of works related to body expression of emotions is still small compared to studies about facial expression.

Analysing some of the few works that have studied human body expressiveness, it is possible to recognize two different methodologies to create the data base of movements to be assessed during the experiments. Applying a first methodology, human actors (either professionals, or amateurs) are asked to walk straight from point A to point B conveying specific emotions~\cite{Dael2012,Meijer1989,Wallboot1998}. Each trial is recorded and later shown to each subjects that have to classify all the sequences. The second methodology uses virtual agents~\cite{Roether2009,Venture2014} to generate the very same set up for the experiments; the agent's movements are generated from the data recorded from human actors.

The work done by Wallboot~\cite{Wallboot1998} has been used as a reference by others. It addressed the question: \textit{Are specific body postures indicative of emotion or are they only used to indicate the intensity of the emotion?} To answer this question, he recorded 224 videos for joy, happiness, sadness, despair, fear, terror, coldness anger, hot anger, disgust, contempt, shame, guilt, pride, and boredom, and showed them to the subjects. His results reaffirm that movement and body postures are indicative of intensity for some emotions. But at the same time, these two characteristics seem to be enough to identify other emotions.

Complementary studies using virtual agents have been performed by  Kluwer and collaborators~\cite{Kluwer2004}, who studied the contribution of postures and angle view in the interpretation of emotions. They generated 176 static positions for happiness, anger, disgust, fear, sadness and surprise. Each image was later rendered from three different angles (front, left, and above and behind left shoulder) producing a total of 528 images. All the participants were exposed to all 528 images and were asked to label the image with the emotions that best represent it. Their results show that five out of six emotions were quite well recognized independently from the angle of view., while disgust was for some postures confused with fear.

The main drawback of all these projects is the use of video recorded sequences, which misses the impact related to a complete physical experience. For instance, it is clearly different facing an angry robot really rushing against us, or looking at a video where this happens.

\subsection{Robotic Studies}
%%%%%%%%%%%%%%%%%%%%%%

As a direct consequence of the abundance of works in face elicitation in humans, most of the works done in Human-Robot Interaction (HRI) have focused also on faces. One of the most well-known expressive robots is Kismet~\cite{Breazeal2002}, a robotic face able to interact with people and to show emotions. The face had enough degrees of freedom to portray the basic emotions suggested by Ekman~\cite{Ekman2004} (happiness, surprise, anger, disgust, fear, and sadness), plus interest. 
Despite the complex system behind Kismet, the emotion's projection evaluation was done using videos with a very limited number of participants. Similar approach was followed by Li and Chignell~\cite{Li2011}, who used videos of a teddy bear robot to study the contribution of arms and head movement to express emotions. In same direction Destephe and collaborators~\cite{Destephe2013b,Destephe2013} studied the attribution of emotion to a robot's gait using a virtual representation of the platform WABIAN-2R. More recently Knight and Simmons~\cite{knight2016} used Keepon and NAO platforms to study the possibility to project inner states with just head movements. Although use of videos has the advantage to cover a major number of participants, the lack of interaction with the real platform make these works lose the impact that a robot could generate on the participants.

%%%%%%%%%%%%%%%%%%%%%%%%
The use of real platforms to study emotion projection could be dived in two lines: using anthropomorphic and non-anthropomorphic platforms. In the first case, these works are characterized by the use of cue positions to project desire emotions~\cite{NAO2013}. In some cases, special attention has been taken to determine head's angle and arms position contribution in specific emotions~\cite{Brown2014}. 
Nevertheless, current humanoid platforms cannot generate smooth gaits, which limit the study of body movements. To overcome this limitation, some researchers have reduce the human appearance (e.g. eliminating limps and facial expressions)to increase platform mobility and study new mechanisms to project emotions~\cite{Arras2012}. Pushing forward the reduction on anthropomorphic features, Saerbeck and Christoph used a Roomba platform to study the contribution of curvature in a trajectory to conceive emotional states~\cite{Saerbeck2010}. Similarly Lourens and Barakova~\cite{BarakovaL10} implemented a set of behaviors to determine the emotion  perceived from diverse movements, which were selected from the  work done by Camurri et al.~\cite{pop00002}. Continuing with her research on how robotics' behaviours are interpreted by people, Barakova and collaborators~\cite{Barakova2013} created a closet in which lights could be manipulated to convey pre-defined behaviours. The robot's behaviours were defined using the Interpersonal Behaviour Circle (ICB)~\cite{Leary57}, which is based on two dimensions (dominance-submission and hate-love). Their findings suggest that electronic systems can elicit a type of reactions different from the one expected by theories of interpersonal communication.

@@

Due to the popularity that quadrocoptors have received in the last years, Sharma and collaborators~\cite{Sharma2013} used a quadrotor to study how different Laban's effort~\cite{Laban1968} parameters could impact on the perception of affection. A professional Laban certified actor was asked to generate 16 different paths, for each one changing one of the four Laban's parameters (space, weight, time, and flow). Each generated path was recorded using the Vicon motion-tracking system. Continuing with the use of quadrocoptors, Cauchard and collaborators~\cite{Cauchard2016} studied how flight paths could project personal traits and emotional attributes. All these works present a very nice starting to point to identify features and values that could be used to project emotions in robotics, which could help in coordinating humans and robots~\cite{Novika2015}. However, all of these works not give a price guideline to elicit precise emotions, which from our previous case studies could lead to the misinterpretation of movements emotions~\cite{Angel2016}. For example people could confuse an implementation of \textit{Happiness} with \textit{Anger}.

>>>>

Most of the works done in Human-Robot Interaction (HRI) have focused on faces, due to the abundance of works in face elicitation in humans. One of the most well-known expressive robots is Kismet~\cite{Breazeal2002}, a robotic face able to interact with people and to show emotions. The face had enough degrees of freedom to portray the basic emotions suggested by Ekman~\cite{Ekman2004} (\textit{Happiness}, \textit{Surprise}, \textit{Anger}, \textit{Disgust}, \textit{Fear}, and \textit{Sadness}), plus interest. 
Despite the complex system behind Kismet, the emotion's projection evaluation was done using videos with a very limited number of participants. Similar approach was followed by Li and Chignell~\cite{Li2011}, who used videos of a teddy bear robot to study the contribution of arms and head movement to express emotions. In same direction Destephe and collaborators~\cite{Destephe2013} studied the attribution of emotion to a robot's gait using a virtual representation of the platform WABIAN-2R. Knight and Simmons~\cite{knight2016} used two platforms (i.e. Keepon and NAO) with different degrees of freedom to study the possibility to project inner states with just head movements. Although the use of videos has the advantage to cover a major number of participants, they loss the impact that is created from the interaction between the participant and the platform.

%%%%%%%%%%%%%%%%%%%%%%%%
The use of real platforms to study emotion projection could be dissect by the type of platform used. Therefore two tendencies could been observed: use of anthropomorphic and non-anthropomorphic platforms. In the first case, these works are characterized by the use of cue positions to project desire emotions~\cite{NAO2013}. In some cases, special attention has been taken to determine head's angle and arms position contribution in specific emotions~\cite{Brown2014}. 
Nevertheless, current humanoid platforms cannot generate smooth gaits, which limit the study of body movements. To overcome this limitation, some researchers have reduce the human appearance (e.g. eliminating limps and facial expressions)to increase platform mobility and study new mechanisms to project emotions~\cite{Arras2012}. Trying to reduce the anthropomorphism, Saerbeck and Christoph used a Roomba platform to study the contribution of curvature in a trajectory to conceive emotional states~\cite{Saerbeck2010}. Similarly Lourens and Barakova~\cite{BarakovaL10} implemented a set of behaviors to determine the emotion  perceived from diverse movements, which were selected from the  work done by Camurri et al.~\cite{pop00002}. Continuing with her research on how robotics' behaviours are interpreted by people, Barakova and collaborators~\cite{Barakova2013} created a closet in which lights could be manipulated to convey pre-defined behaviours. The robot's behaviours were defined using the Interpersonal Behaviour Circle (ICB)~\cite{Leary57}, which is based on two dimensions (dominance-submission and hate-love). Their findings suggest that electronic systems can elicit a type of reactions different from the one expected by theories of interpersonal communication.

@@

Due to the popularity that quadrocoptors have received in the last years, Sharma and collaborators~\cite{Sharma2013} used a quadrotor to study how different Laban's effort~\cite{Laban1968} parameters could impact on the perception of affection. A professional Laban certified actor was asked to generate 16 different paths, for each one changing one of the four Laban's parameters (space, weight, time, and flow). Each generated path was recorded using the Vicon motion-tracking system. Continuing with the use of quadrocoptors, Cauchard and collaborators~\cite{Cauchard2016} studied how flight paths could project personal traits and emotional attributes. All these works present a very nice starting to point to identify features and values that could be used to project emotions in robotics, which could help in coordinating humans and robots~\cite{Novika2015}. However, these works not give a price guideline to elicit precise emotions. This could lead to implementations that express a not desired emotion~\cite{Angel2016}. For example people could confuse an implementation of \textit{Happiness} with \textit{Anger}.
